---
editor_options:
  markdown:
    wrap: 72
output:
  html_document:
    df_print: paged
---

x--- title: "EDA" output: html_document: default word_document: default
date: "2023-11-28" ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Load packages for data manipulation
library(tidyverse)
library(lubridate)
library(ggplot2)
library(cowplot)
library(dplyr)
library(DT)
library(GGally)
library(cluster)
library(factoextra) # for visualization
library(mice)

# Load data
maindata <- read.csv("maindata.csv")
data_monthly <- read.csv("data_monthly.csv")
```


```{r message=FALSE, warning=FALSE}
#data preprocessing
# Create a monthly date sequence from January 1990 to November 2023.
dates <- seq(ymd("1990-01-01"), ymd("2023-11-01"), by = "1 month")

# Assume maindata has been loaded into the environment. Now, add the date column.
maindata <- maindata %>%
  mutate(date = dates)


# Remove variables with excessive missing values (NA)
maindata <- maindata %>% select(-solarradiation, -uvindex)

# Replace NA values with 0
maindata <- maindata %>%
  mutate(
    Clear = replace(Clear, is.na(Clear), 0),
    Overcast = replace(Overcast, is.na(Overcast), 0),
    Partially.cloudy = replace(Partially.cloudy, is.na(Partially.cloudy), 0),
    Rain..Partially.cloudy = replace(Rain..Partially.cloudy, is.na(Rain..Partially.cloudy), 0),
    Rain..Overcast = replace(Rain..Overcast, is.na(Rain..Overcast), 0),
    Rain = replace(Rain, is.na(Rain), 0)
  )
# Rename
maindata <- maindata %>%
  rename(
    'Temperature' = temp,
    'Humidity' = humidity,
    'Wind Speed' =  windspeed ,
    'Sea Level Pressure' = sealevelpressure,
    'Cloud Cover' = cloudcover,
    'Precipitation' = precip,
    `Partially Cloudy` = Partially.cloudy,
    `Rain & Partially Cloudy` = Rain..Partially.cloudy,
    `Rain & Overcast` = Rain..Overcast
  )
```

```
The preprocessing involves creating a monthly date sequence from January 1990 to November 2023, which is then added to maindata. This dataset includes various weather-related variables like Temperature, Humidity,  Wind Speed , cloud cover, Sea Level Pressure, Precipitationitation and different conditions (Clear, Overcast, Partially cloudy,  Rain & Partially, Rain & Overcast, Rain).
```

## longtype
```{r message=FALSE, warning=FALSE}

# Melt the data into a long format suitable for ggplot2, excluding 'uvindex' and 'solarradiation' due to NaN values.
maindata_long <- maindata %>%
  gather(key = "variable", value = "value", -date) %>%
  filter(!is.nan(value))
```

## Plot the annual mean Temperature
```{r message=FALSE, warning=FALSE}
# Assuming maindata_long is already created with the 'date' and 'value' columns
# and the 'variable' column indicating the type of measurement, such as Temperature.

# First, we will filter out Temperature-related variables only
maindata_Temperature <- maindata_long %>%
  filter(variable %in% c("Temperature")) # Replace with actual Temperature variable names

# Extract the year from the date
maindata_Temperature <- maindata_Temperature %>%
  mutate(year = year(date))

# Assuming maindata_long is already created with the 'date' and 'value' columns
# and the 'variable' column indicating the type of measurement, such as Temperature.

# First, we will filter out Temperature-related variables only
maindata_Temperature <- maindata_long %>%
  filter(variable %in% c("Temperature")) # Replace with actual Temperature variable names

# Extract the year from the date
maindata_Temperature <- maindata_Temperature %>%
  mutate(year = year(date))

# Calculate the annual mean Temperature
annual_mean_Temperature <- maindata_Temperature %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean Temperature
annual_mean_Temperature_plot <- annual_mean_Temperature %>%
  filter(variable == "Temperature") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Temperature",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Degrees Celsius"
  ) +
  NULL

annual_mean_Temperature_plot
```

## Plot the monthly mean Temperature
```{r message=FALSE, warning=FALSE}
# Assuming maindata_long has already been created and includes the 'date' and 'value' columns
# We'll use the 'Temperature' variable for plotting the mean Temperature

# First, extract the year and month from the date
maindata_Temperature <- maindata_long %>%
  filter(variable == "Temperature") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_Temperature <- maindata_Temperature %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Temperature
Temperature_plot <- ggplot(monthly_mean_Temperature, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Mean Temperature",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Degrees Celsius"
  ) +
  facet_wrap(~month) +
  NULL

Temperature_plot
```

## Plot the annual mean Humidity
```{r message=FALSE, warning=FALSE}
maindata_Humidity<- maindata_long %>%
  filter(variable %in% c("Humidity")) 

# Extract the year from the date
maindata_Humidity <- maindata_Humidity %>%
  mutate(year = year(date))

# First, we will filter out Temperature-related variables only
maindata_Humidity <- maindata_long %>%
  filter(variable %in% c("Humidity")) 

# Extract the year from the date
maindata_Humidity <- maindata_Humidity %>%
  mutate(year = year(date))

# Calculate the annual mean Humidity
annual_mean_Humidity <- maindata_Humidity %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean Humidity with the title and subtitle centered
annual_mean_Humidity_plot <- annual_mean_Humidity %>%
  filter(variable == "Humidity") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Humidity",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "%"
  ) +
  NULL

annual_mean_Humidity_plot
```

## Plot the monthly mean Humidity
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_Humidity <- maindata_long %>%
  filter(variable == "Humidity") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_Humidity <- maindata_Humidity %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Temperature
Humidity_plot <- ggplot(monthly_mean_Humidity, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Mean Humidity",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Degrees Celsius"
  ) +
  facet_wrap(~month) +
  NULL

Humidity_plot
```

## Plot the annual mean Wind Speed 
```{r message=FALSE, warning=FALSE}
maindata_WindSpeed <- maindata_long %>%
  filter(variable %in% c("Wind Speed")) 

# Extract the year from the date
maindata_WindSpeed  <- maindata_WindSpeed  %>%
  mutate(year = year(date))

# First, we will filter out Temperature-related variables only
maindata_WindSpeed  <- maindata_long %>%
  filter(variable %in% c("Wind Speed")) 

# Extract the year from the date
maindata_WindSpeed  <- maindata_WindSpeed  %>%
  mutate(year = year(date))

# Calculate the annual mean Humidity
annual_mean_WindSpeed  <- maindata_WindSpeed  %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean  Wind Speed  with the title and subtitle centered
annual_mean_WindSpeed_plot <- annual_mean_WindSpeed  %>%
  filter(variable == "Wind Speed") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Wind Speed",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Kilometers per hour (kph) "
  ) +
  NULL

annual_mean_WindSpeed_plot
```

## Plot the monthly mean  Wind Speed 
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_WindSpeed<- maindata_long %>%
  filter(variable == "Wind Speed") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_WindSpeed  <- maindata_WindSpeed  %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Wind Speed
 WindSpeed_plot <- ggplot(monthly_mean_WindSpeed , aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Mean Wind Speed",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Kilometers per hour (kph) "
  ) +
  facet_wrap(~month) +
  NULL

 WindSpeed_plot
```

## Plot the annual mean Cloud Cover
```{r message=FALSE, warning=FALSE}
maindata_CloudCover<- maindata_long %>%
  filter(variable %in% c("Cloud Cover")) 

# Extract the year from the date
maindata_CloudCover <- maindata_CloudCover %>%
  mutate(year = year(date))

# First, we will filter out Temperature-related variables only
maindata_CloudCover <- maindata_long %>%
  filter(variable %in% c("Cloud Cover")) 

# Extract the year from the date
maindata_CloudCover <- maindata_CloudCover %>%
  mutate(year = year(date))

# Calculate the annual mean Cloud Cover
annual_mean_CloudCover <- maindata_CloudCover %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean Cloud Cover with the title and subtitle centered
annual_mean_CloudCover_plot <- annual_mean_CloudCover %>%
  filter(variable == "Cloud Cover") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Cloud Cover",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "% "
  ) +
  NULL

annual_mean_CloudCover_plot
```

## Plot the monthly mean Cloud Cover
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_CloudCover <- maindata_long %>%
  filter(variable == "Cloud Cover") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_CloudCover <- maindata_CloudCover %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Temperature
CloudCover_plot <- ggplot(monthly_mean_CloudCover, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Mean Cloud Cover",
    subtitle = "Data from 1990 - 2023",
    y = "%"
  ) +
  facet_wrap(~month) +
  NULL

CloudCover_plot
```

## Plot the annual mean Sea Level Pressure
```{r message=FALSE, warning=FALSE}
maindata_SeaLevelPressure<- maindata_long %>%
  filter(variable %in% c("Sea Level Pressure")) 

# Extract the year from the date
maindata_SeaLevelPressure <- maindata_SeaLevelPressure %>%
  mutate(year = year(date))

# First, we will filter out Temperature-related variables only
maindata_SeaLevelPressure <- maindata_long %>%
  filter(variable %in% c("Sea Level Pressure")) 

# Extract the year from the date
maindata_SeaLevelPressure <- maindata_SeaLevelPressure %>%
  mutate(year = year(date))

# Calculate the annual mean Humidity
annual_mean_SeaLevelPressure <- maindata_SeaLevelPressure %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean Sea Level Pressure with the title and subtitle centered
annual_mean_SeaLevelPressure_plot <- annual_mean_SeaLevelPressure %>%
  filter(variable == "Sea Level Pressure") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Sea Level Pressure",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Millibars (mb)"
  ) +
  NULL

annual_mean_SeaLevelPressure_plot
```

## Plot the monthly mean Sea Level Pressure
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_SeaLevelPressure <- maindata_long %>%
  filter(variable == "Sea Level Pressure") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_SeaLevelPressure <- maindata_SeaLevelPressure %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Temperature
SeaLevelPressure_plot <- ggplot(monthly_mean_SeaLevelPressure, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Mean Sea Level Pressure",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Millibars (mb)"
  ) +
  facet_wrap(~month) +
  NULL

SeaLevelPressure_plot
```

## Plot the annual mean Precipitation
```{r message=FALSE, warning=FALSE}
maindata_Precipitation<- maindata_long %>%
  filter(variable %in% c("Precipitation")) 

# Extract the year from the date
maindata_Precipitation <- maindata_Precipitation %>%
  mutate(year = year(date))

# First, we will filter out Temperature-related variables only
maindata_Precipitation <- maindata_long %>%
  filter(variable %in% c("Precipitation")) 

# Extract the year from the date
maindata_Precipitation <- maindata_Precipitation %>%
  mutate(year = year(date))

# Calculate the annual mean Humidity
annual_mean_Precipitation <- maindata_Precipitation %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean Precipitation with the title and subtitle centered
annual_mean_Precipitation_plot <- annual_mean_Precipitation %>%
  filter(variable == "Precipitation") %>% 
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Precipitationitation",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "mm"
  ) +
  NULL

annual_mean_Precipitation_plot
```

## Plot the monthly mean Precipitation
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_Precipitation <- maindata_long %>%
  filter(variable == "Precipitation") %>%
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean Temperature
monthly_mean_Precipitation <- maindata_Precipitation %>%
  group_by(year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot the monthly mean Temperature
Precipitation_plot <- ggplot(monthly_mean_Precipitation, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Precipitationitation",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "mm"
  ) +
  facet_wrap(~month) +
  NULL

Precipitation_plot
```


## Plot the annual mean conditions: 'Clear', 'Overcast', 'Partially cloudy','Rain & Partially cloudy', 'Rain & Overcast', 'Rain'
```{r message=FALSE, warning=FALSE}
# First, we will filter out condition-related variables only
maindata_condition <- maindata_long %>% 
  filter(variable %in% 
           c("Clear","Overcast","Partially Cloudy", 
             "Rain & Partially Cloudy", "Rain & Overcast", "Rain")) 


# Extract the year from the date
maindata_condition <- maindata_condition %>%
  mutate(year = year(date))

# Calculate the annual mean frequency
annual_mean_condition <- maindata_condition %>%
  group_by(year, variable) %>%
  summarise(year_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Plot the annual mean frequency with the title and subtitle centered
annual_mean_condition_plot <- annual_mean_condition %>%
  ggplot(aes(x = year, y = year_mean)) +
  geom_point() +
  geom_line() +
  geom_smooth(method = "loess") +
  facet_wrap(~variable) + 
  theme(
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    plot.subtitle = element_text(hjust = 0.5) 
  ) +
  labs(
    title = "Annual Mean Frequency of Weather Condition",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Count"
  ) +
  NULL

annual_mean_condition_plot
```


## Plot the monthly mean conditions:
```{r message=FALSE, warning=FALSE}
# First, extract the year and month from the date
maindata_condition <- maindata_long %>% 
  filter(variable %in% 
           c("Clear","Overcast","Partially Cloudy", 
             "Rain & Partially Cloudy", "Rain & Overcast", "Rain")) %>% 
  mutate(year = year(date), month = month(date, label = TRUE))

# Calculate the monthly mean frequency
monthly_mean_condition <- maindata_condition %>%
  group_by(variable, year, month) %>%
  summarise(month_mean = mean(value, na.rm = TRUE)) %>%
  ungroup()

# Define a colour scale for the months
my_colour <- scales::hue_pal()(12)

# Plot 
condition_plot <- ggplot(monthly_mean_condition, aes(x = year, y = month_mean, colour = month)) +
  geom_point(size = 0.5) +
  geom_smooth(method = "loess") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  ) +
  scale_color_manual(values = my_colour) +
  labs(
    title = "Monthly Frequency of Weather Condition",
    subtitle = "Data from Jan. 1990 to Nov. 2023",
    y = "Count"
  ) +
  facet_grid(variable~month) +
  NULL

condition_plot
```


## Boxplot: Temperature
```{r message=FALSE, warning=FALSE}
Temperature_data <- maindata_long %>%
  filter(variable == "Temperature") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
Temperature_boxplot <- ggplot(Temperature_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Temperature",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Degrees Celsius") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Temperature_boxplot)
```

## Boxplot: Temperature with outlier
```{r message=FALSE, warning=FALSE}

Temperature_data <- maindata_long %>%
  filter(variable == "Temperature") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <- Temperature_data %>%
  group_by(month) %>%
  summarise(lower = quantile(value, 0.25) - 1.5 * IQR(value),
            upper = quantile(value, 0.75) + 1.5 * IQR(value)) %>%
  left_join(Temperature_data, by = "month") %>%
  filter(value < lower | value > upper)

Temperature_boxplot_outlier <- ggplot(Temperature_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Temperature",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Degrees Celsius") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Temperature_boxplot_outlier)
```

## Boxplot: Humidity
```{r message=FALSE, warning=FALSE}

Humidity_data <- maindata_long %>%
  filter(variable == "Humidity") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
Humidity_boxplot <- ggplot(Humidity_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Humidity",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Humidity_boxplot)
```
## Boxplot: Humidity with outlier
```{r message=FALSE, warning=FALSE}

Humidity_data <- maindata_long %>%
  filter(variable == "Humidity") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <- Humidity_data %>%
  group_by(month) %>%
  summarise(lower = quantile(value, 0.25) - 1.5 * IQR(value),
            upper = quantile(value, 0.75) + 1.5 * IQR(value)) %>%
  left_join(Humidity_data, by = "month") %>%
  filter(value < lower | value > upper)

Humidity_boxplot_outlier <- ggplot(Humidity_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Humidity",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Humidity_boxplot_outlier)
```

## Boxplot:  Wind Speed 
```{r message=FALSE, warning=FALSE}

 WindSpeed_data <- maindata_long %>%
  filter(variable == "Wind Speed") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
 WindSpeed_boxplot <- ggplot( WindSpeed_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Wind Speed",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Kilometers per hour (kph)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(WindSpeed_boxplot)
```

## Boxplot:  Wind Speed  with outlier
```{r message=FALSE, warning=FALSE}

 WindSpeed_data <- maindata_long %>%
  filter(variable == "Wind Speed") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <-  WindSpeed_data %>%
  group_by(month) %>%
  summarise(lower = quantile(value, 0.25) - 1.5 * IQR(value),
            upper = quantile(value, 0.75) + 1.5 * IQR(value)) %>%
  left_join( WindSpeed_data, by = "month") %>%
  filter(value < lower | value > upper)

 WindSpeed_boxplot_outlier <- ggplot( WindSpeed_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Wind Speed",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Kilometers per hour (kph)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(WindSpeed_boxplot_outlier)
```

## Boxplot: Cloud Cover
```{r message=FALSE, warning=FALSE}
CloudCover_data <- maindata_long %>%
  filter(variable == "Cloud Cover") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
CloudCover_boxplot <- ggplot(CloudCover_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Cloud Cover",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(CloudCover_boxplot)
```

## Boxplot: Cloud Cover with outlier
```{r message=FALSE, warning=FALSE}

CloudCover_data <- maindata_long %>%
  filter(variable == "Cloud Cover") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <- CloudCover_data %>%
  group_by(month) %>%
  summarise(lower = quantile(value, 0.25) - 1.5 * IQR(value),
            upper = quantile(value, 0.75) + 1.5 * IQR(value)) %>%
  left_join(CloudCover_data, by = "month") %>%
  filter(value < lower | value > upper)

CloudCover_boxplot_outlier <- ggplot(CloudCover_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Cloud Cover",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "%") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(CloudCover_boxplot_outlier)
```

## Boxplot: Sea Level Pressure
```{r message=FALSE, warning=FALSE}
SeaLevelPressure_data <- maindata_long %>%
  filter(variable == "Sea Level Pressure") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
SeaLevelPressure_boxplot <- ggplot(SeaLevelPressure_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Sea Level Pressure",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Millibars (mb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(SeaLevelPressure_boxplot)
```

## Boxplot: Sea Level Pressure with outlier
```{r message=FALSE, warning=FALSE}

SeaLevelPressure_data <- maindata_long %>%
  filter(variable == "Sea Level Pressure") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <- SeaLevelPressure_data %>%
  group_by(month) %>%
  summarise(lower = quantile(value, 0.25) - 1.5 * IQR(value),
            upper = quantile(value, 0.75) + 1.5 * IQR(value)) %>%
  left_join(SeaLevelPressure_data, by = "month") %>%
  filter(value < lower | value > upper)

SeaLevelPressure_boxplot_outlier <- ggplot(SeaLevelPressure_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Sea Level Pressure",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "Millibars (mb)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(SeaLevelPressure_boxplot_outlier)
```

## Boxplot: Precipitation
```{r message=FALSE, warning=FALSE}
Precipitation_data <- maindata_long %>%
  filter(variable == "Precipitation") %>%
  mutate(month = factor(month(date, label = TRUE)))  # Converting the date to a month factor

# Creating the boxplot for each month
Precipitation_boxplot <- ggplot(Precipitation_data, aes(x = month, y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Boxplots of Precipitation",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "mm") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Precipitation_boxplot)
```

## Boxplot: Precipitation with Precipitation
```{r message=FALSE, warning=FALSE}

Precipitation_data <- maindata_long %>%
  filter(variable == "Precipitation") %>%
  mutate(year = as.character(year(date)),
         month = factor(month(date, label = TRUE)))

outliers <- Precipitation_data %>%
  group_by(month) %>%
  summarise(
    lower = quantile(value, 0.25, na.rm = TRUE) - 1.5 * IQR(value, na.rm = TRUE),
    upper = quantile(value, 0.75, na.rm = TRUE) + 1.5 * IQR(value, na.rm = TRUE)
  ) %>%
  left_join(Precipitation_data, by = "month") %>%
  filter(value < lower | value > upper)


Precipitation_boxplot_outlier <- ggplot(Precipitation_data, aes(x = month, y = value)) +
  geom_boxplot() +
  geom_text(data = outliers, aes(label = year), vjust = -0.5) +
  labs(title = "Monthly Boxplots of Precipitation",
       subtitle = "Data from Jan. 1990 to Nov. 2023",
       x = "Month",
       y = "mm") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))

print(Precipitation_boxplot_outlier)
```


# Yearly Basic Statistics Table (excluding the condition variables)
# feture: Temperature, Humidity,  Wind Speed , Cloud Cover, Sea Level Pressure, Precipitation
# If you want to summarize the data by month instead of by year, you can add 'group_by(year, month)'

```{r message=FALSE, warning=FALSE}
library(DT)
maindata %>% 
    mutate(year = year(date), month = month(date, label = TRUE)) -> data_summary

data_summary %>%
  select('year', 'Temperature', 'Humidity', 'Wind Speed', 'Cloud Cover', 'Sea Level Pressure', 'Precipitation') %>% 
  group_by(year) %>%
  summarise(across(everything(), 
                   list(min = ~min(., na.rm = TRUE), 
                       max = ~max(., na.rm = TRUE), 
                       mean = ~mean(., na.rm = TRUE)))) %>%
  mutate_all(~round(., 3)) -> datatable1

datatable(datatable1, options = list(autoWidth = FALSE, scrollX = TRUE))

```

# Yearly Basic Statistics Table (only the condition variables)
# If you want to summarize the data by month instead of by year, you can add 'group_by(year, month)'
```{r message=FALSE, warning=FALSE}
data_summary %>% 
    select(Clear:month) %>%
    group_by(year) %>%
    summarise_at(vars(Clear:Rain), 
                      list(min = ~min(., na.rm = TRUE), 
                           max = ~max(., na.rm = TRUE), 
                           mean = ~mean(., na.rm = TRUE))) %>%
  mutate_all(~ifelse(is.finite(.), ., NA)) %>% 
  mutate_all(~round(., 3)) -> datatable2

datatable(datatable2, options = list(autoWidth = FALSE, scrollX = TRUE))

```

# Yearly Basic Statistics Table
# If you want to summarize the data by month instead of by year, you can add 'group_by(year, month)'
```{r message=FALSE, warning=FALSE}

# Joining the two tables on the 'year' column
final_table <- left_join(datatable1, datatable2, by = "year")

# Displaying the combined table
datatable(final_table, options = list(autoWidth = FALSE, scrollX = TRUE))

```


#Excluding the variables 'Clear', 'Overcast', 'Partially Cloudy', 'Rain & Partially Cloudy', 'Rain & Overcast', 'Rain', as they make the data look too messy. Should I add them back?
# If you want to summarize the data by month instead of by year, you can add 'group_by(year, month)'
```{r message=FALSE, warning=FALSE}
# heatmap(Excluding the 'Clear', 'Overcast', 'Partially Cloudy','Rain & Partially Cloudy', 'Rain & Overcast', 'Rain')
data_summary %>%
  select('Temperature', 'Humidity', 'Wind Speed', 'Cloud Cover', 'Sea Level Pressure', 'Precipitation') -> cordata

library(GGally)
ggpairs(cordata)

cordata %>%
  cor(use = "complete.obs") -> cor_matrix

heatmap(cor_matrix, Colv=NA, Rowv = NA, scale = "column") 

library(corrplot)
corrplot(cor_matrix, method = 'color', type = 'lower', order = 'hclust',
         tl.col = 'black', addCoef.col = 'black', number.cex = 0.6,
         cl.ratio = 0.2, tl.srt = 45, col = COL2('RdBu', 10))


## leave blank on non-significant coefficient
## add significant correlation coefficients
testRes = cor.mtest(cordata, conf.level = 0.95)

corrplot(cor_matrix, p.mat = testRes$p, 
         method = 'color', type = 'lower', insig='blank',
         tl.col = 'black', addCoef.col = 'black', number.cex = 0.6,
         order = 'AOE', diag=FALSE)
```

